---
title: "AIED_Prediction_Assignment1"
author: "20201564 김성현"
date: "`r Sys.Date()`"
output: word_document
---

## Set the working directory and read in the dataset (`PISA2018MS_KOR_BQ.rdata`)

```{r readin}
infile <- outfile <-  "C:/aied"
setwd(infile) 
getwd()
load("PISA2018MS_KOR_BQ.Rdata")
class(PISA2018MS_KOR)
```

## Prediction task using the linear regression

### Check the dimension of the dataset
```{r dim}
dim(PISA2018MS_KOR)
```

### Compute the mean and sd of PV1MATH variable (outcome variable)
```{r outcome}
mean(PISA2018MS_KOR$PV1MATH)
sd(PISA2018MS_KOR$PV1MATH)
```

### Fit the following three prediction models and return the summary of each model: 
* M1: PV1MATH by EMOSUPS
* M2: PV1MATH by explanatory variable: EMOSUPS,  ST004D01T (gender)
* M3: PV1MATH by explanatory variables: EMOSUPS, ST004D01T (gender), interaction between EMOSUPS and ST004D01T (gender)
```{r linear_reg}
M1 <- lm(PV1MATH ~ EMOSUPS, data=PISA2018MS_KOR)
summary(M1)
M2 <- lm(PV1MATH ~ as.factor(ST004D01T) + EMOSUPS, data = PISA2018MS_KOR)
summary(M2)
M3 <- lm(PV1MATH ~ as.factor(ST004D01T) +  EMOSUPS + as.factor(ST004D01T):EMOSUPS, data = PISA2018MS_KOR)
summary(M3)
```

### Interpret the coefficients of M2 and M3
#### M2 Results
* intercept = 522.432: 
* ST004D01T =  7.162: 
* EMOSUPS   = 15.699: 

#### M3 Results
* intercept = 522.544: 
* ST004D01T =  6.937:
* EMOSUPS   =  14.548:
* as.factor(ST004D01T) =  2.190:
* Interaction term shows that One unit increase in EMOSUPS is associated with 14.548 higher mathematics score, on average, for girls, and 16.738 (14.548+2.190) higher mathematics score, on average, for boys.

# Check the assumptions of M2 using the `plot` function
```{r assumption}
plot(M2)
```

