---
title: "AIED_Classification_Assignment2"
author: "Hyo Jeong Shin"
date: "`r Sys.Date()`"
output: word_document
---

### Set the working directory and read in the dataset (`PISA2018MS_KOR_BQ.rdata`)

```{r readin}
infile <- outfile <-  "/Users/hyojeong/Library/Mobile Documents/com~apple~CloudDocs/Teaching/2023_1/UnderGrad_AIED"
setwd(infile)
getwd()
load("PISA2018MS_KOR_BQ.Rdata")
```

### Classification using the logistic regression

#### Make the table of the outcome variable `EC154Q02IA`
* EC154Q02IA: attending additional instruction in mathematics
* Make sure to to include if there are any missing cases (`useNA='always'`)

```{r NAcheck}
table(PISA2018MS_KOR$EC154Q02IA, useNA='always')
```

#### Treatment of outcome variable
* Remove the missing cases of the outcome variable 
* Check the dimension
* Recode 1 -> 0 & 2 -> 1
* Create the table again to check if the recoding is done successfully
```{r outcome}
PISA2018MS_KOR <- PISA2018MS_KOR[!is.na(PISA2018MS_KOR$EC154Q02IA),]
dim(PISA2018MS_KOR) 
PISA2018MS_KOR$EC154Q02IA <- PISA2018MS_KOR$EC154Q02IA - 1
table(PISA2018MS_KOR$EC154Q02IA, useNA='always')
```

###  Three ways to compute proportions: outcome variable by gender 
#### Create three types of tables 
```{r proportions}
contab <- table(Math = PISA2018MS_KOR$EC154Q02IA, Gender = PISA2018MS_KOR$ST004D01T)
prop.table(contab) # out of all students (N=6429 excluding missing cases)
prop.table(contab, margin = 1)
prop.table(contab, margin = 2)
```

#### Interpret the proportions of each table

* Out of total observations, 30.4% are male who do not attend additional instruction in math, while 21.2% are male who attend additional instruction in math. Similarly, 29.4% are female who do not attend additional instruction in math, while 19.1% are female who attend additional instruction in math.
* Among the students who do not attend additional instruction in math (conditional on math == 2), 50.9% are male students while 49.1% are female students. Among the students who attend additional instruction in math (conditional on math == 1), 52.6% are male students while 47.4% are female students.
* Among female students (conditional on gender == 1), 39.4% attend additional instruction in math, while 60.6% do not attend. Among male students (conditional on gender == 2), 41.0% attend additional instruction in math, while 59.0% do not attend.

### Logistic regression
#### Fit the following three logistic regressions (no interpretation is required)
* M1: EC154Q02IA by ST004D01T (gender)
* M2: EC154Q02IA by explanatory variable: PV1MATH,  ST004D01T (gender)
* M3: EC154Q02IA by explanatory variables: PV1MATH, ST004D01T (gender), interaction between PV1MATH and ST004D01T (gender)

```{r logit_reg}
M1 <- glm(EC154Q02IA ~ as.factor(ST004D01T), data = PISA2018MS_KOR, family = "binomial")
summary(M1)
M2 <- glm(EC154Q02IA ~ as.factor(ST004D01T)+ PV1MATH, data = PISA2018MS_KOR, family = "binomial")
summary(M2)
M3 <- glm(EC154Q02IA ~ as.factor(ST004D01T)+ PV1MATH + as.factor(ST004D01T):PV1MATH, data = PISA2018MS_KOR, family = "binomial")
summary(M3)
```

### [BONUS] Model Evaluation (5 points)
#### Predict the probabilities and values (either 0 or 1) from each model
* Use `get_logistic_pred` function in the lecture notes
* You need to generate three sets of predicted values based on each model (M1, M2, M3)
```{r logit_pred}
get_logistic_pred = function(mod, data, res = "y", pos = 1, neg = 0, cut = 0.5) {
  probs = predict(mod, newdata = data, type = "response")
  ifelse(probs >= cut, pos, neg)
}
M1_pred <- get_logistic_pred(M1, data = PISA2018MS_KOR, res = "EC154Q02IA", cut = 0.5)
M2_pred <- get_logistic_pred(M2, data = PISA2018MS_KOR, res = "EC154Q02IA", cut = 0.5)
M3_pred <- get_logistic_pred(M3, data = PISA2018MS_KOR, res = "EC154Q02IA", cut = 0.5)
```

#### Evaluate models (M2 & M3) based on the following quantities
* Report Accuracy, Specificity, Subjectivity, F1 score from M2 & M3
* Use `confusion matrix`
```{r ROC}
M2_tab <- table(pred = M2_pred, actual = PISA2018MS_KOR$EC154Q02IA)
M3_tab <- table(pred = M3_pred, actual = PISA2018MS_KOR$EC154Q02IA)

library(caret)
confusionMatrix(M2_tab, positive = "1", mode = "everything")
confusionMatrix(M3_tab, positive = "1", mode = "everything")
```
