---
title: "AIED_Classification_Assignment2"
author: "20201564 김성현"
date: "`r Sys.Date()`"
output: word_document
---

### Set the working directory and read in the dataset (`PISA2018MS_KOR_BQ.rdata`)

```{r readin}
infile <- outfile <-  "C:/aied"
setwd(infile) 
getwd()
load("PISA2018MS_KOR_BQ.Rdata")
class(PISA2018MS_KOR)
```

### Classification using the logistic regression

#### Make the table of the outcome variable `EC154Q02IA`
* EC154Q02IA: attending additional instruction in mathematics
* Make sure to to include if there are any missing cases (`useNA='always'`)

```{r NAcheck}
table(PISA2018MS_KOR$EC154Q02IA, useNA='always')
```

#### Treatment of outcome variable
* Remove the missing cases of the outcome variable 
* Check the dimension
* Recode 1 -> 0 & 2 -> 1
* Create the table again to check if the recoding is done successfully
```{r outcome}
PISA2018MS_KOR <- PISA2018MS_KOR[!is.na(PISA2018MS_KOR$EC154Q02IA),]
dim(PISA2018MS_KOR)
PISA2018MS_KOR$EC154Q02IA <- PISA2018MS_KOR$EC154Q02IA - 1 # recode 1 -> 0 & 2 -> 1

table(PISA2018MS_KOR$EC154Q02IA, useNA='always')
contab <- table(math = PISA2018MS_KOR$EC154Q02IA, gender = PISA2018MS_KOR$ST004D01T)
contab_margins <- addmargins(contab)
contab_margins
```

###  Three ways to compute proportions: outcome variable by gender 
#### Create three types of tables 
```{r proportions}
prop.table(contab)
prop.table(contab, margin = 1)
prop.table(contab, margin = 2)

```

#### Interpret the proportions of each table

* Out of total observations, 30.4% are male who do not attending additional instruction in mathematics while 21.2% are male who do attending additional instruction in mathematics. 29.3% are female who do not attending additional instruction in mathematics while 19.1% are female who do attending additional instruction in mathematics.
* Among the students who do not attend additional instruction in math (conditional on math == 1), 49.1% are female students and 50.9% are male students
* Among female students (conditional on gender == 1), 39.4% do attending additional instruction in mathematics while 60.6% do not attending additional instruction in mathematics.

### Logistic regression
#### Fit the following three logistic regressions (no interpretation is required)
* M1: EC154Q02IA by ST004D01T (gender)
* M2: EC154Q02IA by explanatory variable: PV1MATH,  ST004D01T (gender)
* M3: EC154Q02IA by explanatory variables: PV1MATH, ST004D01T (gender), interaction between PV1MATH and ST004D01T (gender)

```{r logit_reg}
M1 <- glm(EC154Q02IA ~ as.factor(ST004D01T), data = PISA2018MS_KOR, family = "binomial")
summary(M1)
M2 <- glm(EC154Q02IA ~ as.factor(ST004D01T) + PV1MATH, data = PISA2018MS_KOR, family = "binomial")
summary(M2)
M3 <- glm(EC154Q02IA ~ as.factor(ST004D01T) + PV1MATH + as.factor(ST004D01T):PV1MATH, data = PISA2018MS_KOR, family = "binomial")
summary(M3)
```


### [BONUS] Model Evaluation (5 points)
#### Predict the probabilities and values (either 0 or 1) from each model
* Keep `get_logistic_pred` function 
* You need to generate three sets of predicted values based on each model (M1, M2, M3)
```{r logit_pred}
get_logistic_pred = function(mod, data, res = "y", pos = 1, neg = 0, cut = 0.5) {
  probs = predict(mod, newdata = data, type = "response")
  ifelse(probs >= cut, pos, neg)
}
pred_M1 <- get_logistic_pred(M1, data=PISA2018MS_KOR, res="EC154Q02IA", cut = 0.5)
tab_M1 <- table(pred=pred_M1, actual=PISA2018MS_KOR$EC154Q02IA)
pred_M2 <- get_logistic_pred(M2, data=PISA2018MS_KOR, res="EC154Q02IA", cut = 0.5)
tab_M2 <- table(pred=pred_M2, actual=PISA2018MS_KOR$EC154Q02IA)
pred_M3 <- get_logistic_pred(M3, data=PISA2018MS_KOR, res="EC154Q02IA", cut = 0.5)
tab_M3 <- table(pred=pred_M3, actual=PISA2018MS_KOR$EC154Q02IA)
```

#### Evaluate models (M2 & M3) based on the following quantities
* Report Accuracy, Specificity, Subjectivity, F1 score from M2 & M3
* Use `confusion matrix`
```{r confmat}
#install.packages("caret")
library(caret)
confusionMatrix(tab_M2, mode='everything', positive="1")
confusionMatrix(tab_M3,  mode='everything',positive="1")

```