---
title: "AIED_Prediction_Assignment1"
author: "Hyo Jeong Shin"
date: "`r Sys.Date()`"
output: word_document
---

## Set the working directory and read in the dataset (`PISA2018MS_KOR_BQ.rdata`)

```{r readin}
infile <- outfile <-  "/Users/hyojeong/Library/Mobile Documents/com~apple~CloudDocs/Teaching/2023_1/UnderGrad_AIED"
setwd(infile)
getwd()
load("PISA2018MS_KOR_BQ.Rdata")
class(PISA2018MS_KOR)
```

## Prediction task using the linear regression

### Check the dimension of the dataset
```{r dim}
dim(PISA2018MS_KOR)
```

### Compute the mean and sd of PV1MATH variable (outcome variable)
```{r outcome}
mean(PISA2018MS_KOR$PV1MATH)
sd(PISA2018MS_KOR$PV1MATH)
```

### Fit the following three prediction models and return the summary of each model: 
* M1: PV1MATH by EMOSUPS
* M2: PV1MATH by explanatory variable: EMOSUPS,  ST004D01T (gender)
* M3: PV1MATH by explanatory variables: EMOSUPS, ST004D01T (gender), interaction between EMOSUPS and ST004D01T (gender)
```{r linear_reg}
M1 <- lm(PV1MATH ~ EMOSUPS, data = PISA2018MS_KOR)
M2 <- lm(PV1MATH ~ EMOSUPS + as.factor(ST004D01T), data = PISA2018MS_KOR)
M3 <- lm(PV1MATH ~ EMOSUPS + as.factor(ST004D01T) + as.factor(ST004D01T):EMOSUPS, data = PISA2018MS_KOR)
```

### Interpret the coefficients of M2 and M3
#### M2 Results
* intercept = 522.432 : 522.4 is the expected MATH score for girls when EMOSUPS is zero
* ST004D01T = 7.162  : MATH score is 7.16 higher for boys, on average, controlling for the EMOSUPS 
* EMOSUPS   = 15.699 : one unit increase in EMOSUPS is associated with 15.699 higher MATH score, on average, for the given gender (i.e., same EMOSUPS effect for boys and girls)


#### M3 Results
* intercept = 522.544: 522.5 is the expected MATH score for girls when EMOSUPS is zero
* ST004D01T = 6.937 : MATH score is 6.94 higher for boys, on average, when EMOSUPS is zero
* EMOSUPS   = 14.548: one unit increase in EMOSUPS is associated with 14.55 higher MATH score, on average, for girls
* as.factor(ST004D01T):EMOSUPS = 2.190  : one unit increase in EMOSUPS is associated with 16.738 (14.548 + 2.190) higher MATH score, on average, for boys
* Interaction term that is not statistically significant indicates that the effect of EMOSUPS is not statistically different for boys and girls 

# Check the assumptions of M2 using the `plot` function
```{r assumption}
plot(M2)
```

Normal Q-Q plot and the residual vs. fitted plot show that the errors follow the normal distribution and that the errors appear constant. 